---
title: "Classification and GLM"
author: "Pablo X Zumba"
date: "`r Sys.Date()`"
output: word_document
---
## Preprocessing
```{r}
rm(list=ls())
df=rio::import("Lung_cancer.xlsx", sheet="Sheet1")
colnames(df)=tolower(make.names(colnames(df)))
#str(df)
attach(df)
```
## Exploring Data
```{r}
table(treatment)# 69 had standard treatment and 68 test treatment. Pretty balanced.
table(cell_type)#35 Squamous, 48 Small cell, 27 adeno, and 27 large cell type.
hist(survival_in_days, breaks=30)#Non normally distributed, right skewed. Most did not survive over 400 days.
table(status) # 128 dead and 9 censored. This is the EVENT.
hist(karnofsky_score, breaks = 12) #Kinda bimodal. Most observations are between 50 and 80 score.
hist(months_from_diagnosis,breaks=30) #Non normally distributed, right skewed. Most obs. are below 20 months.
hist(age_in_years, breaks = 20)#Most obs are between 60 and 70 year old.
table(prior_chemotherapy) # 97 did not have prior kimo and 40 did. 
```
# Kaplan-Meier non-parametric analysis
1. We would like to see Kaplan-Meier survival graphs for patients with the test vs standard treatment. Use this data to assess:
-**What are the probability that the patient will survive for 1 year (365 days) and 6 months (183 days) on the standard treatment vs the test treatment?**
-**What is the mean number of days where a patient can be expected to survive if they are on the standard vs the test treatment?**
Since we want to compare test vs standart treatment, we have to use Kaplan-Meier non-parametric analysis by treatment, using survival in days as Time and status as event.
```{r}
library(tidyverse)
library(survival)
km_treatment <- survfit(Surv(survival_in_days, status) ~ treatment, data = df)
km_treatment
#64 of 69 indivuals with standard treatmend died.
#64 of 68 indivuals with test treatmend died.
summary(km_treatment)
#plot(km_treatment, xlab="Time", ylab="Survival Probability")
#install.packages("survminer")
library("survminer")
ggsurvplot(
  km_treatment,
  fun = "pct",
  linetype = "strata", # Change line type by groups
  pval = FALSE, 
  pval.method = TRUE,
  conf.int = TRUE,
  risk.table = TRUE,
  fontsize = 3, # used in risk table
  surv.median.line = "hv", # median horizontal and vertical ref lines
  ggtheme = theme_light(),
  palette = c("#E7B800", "#2E9FDF"),
  title = "Kaplan-Meier Survival Function Estimate"
)
summary(km_treatment, time = 365)
km_treatment %>% broom::tidy()
summary(km_treatment, time = 183)
```
# What is the average number of days where a patient can be expected to survive if they are on the standard vs the test treatment?
Assuming the patients are alive a the end of the time-frame window we can subset the data depending on the status and treatment.
```{r}
alive_standart_treat = subset(df, status == 0 & treatment == 1) #Status=censored and treatment=standard
alive_test_treat = subset(df, status == 0 & treatment == 2) #Status=censored and treatment=standard
mean(alive_standart_treat$survival_in_days) # = 105.4 days
mean(alive_test_treat$survival_in_days)     # = 126 days
log_standard_days=log(mean(alive_standart_treat$survival_in_days))
log_test_days=log(mean(alive_test_treat$survival_in_days))
hist(log(survival_in_days))
abline(v=log_standard_days, col="blue")
abline(v=log_test_days, col="red")
```

# 2. Comparison of different semi-parametric and parametric models.
```{r}
# Kaplan-Meier non-parametric analysis by group
kaplan_meier <- survfit(Surv(survival_in_days, status) ~ treatment, data = df)
summary(kaplan_meier)
#plot(kaplan_meier, xlab="Survival in days", ylab="Survival Probability")

# Cox proportional hazard model - coefficients and hazard rates
cox <- coxph(Surv(survival_in_days, status) ~ treatment + months_from_diagnosis + age_in_years + prior_chemotherapy + karnofsky_score + cell_type, method="breslow", data = df)
#summary(cox)

# Exponential, Weibull, and log-logistic parametric model coefficients
exponential <- survreg(Surv(survival_in_days, status) ~ treatment + months_from_diagnosis + age_in_years + prior_chemotherapy + karnofsky_score + cell_type, dist="exponential", data = df)
#summary(exponential)

weibull <- survreg(Surv(survival_in_days, status) ~ treatment + months_from_diagnosis + age_in_years + prior_chemotherapy + karnofsky_score + cell_type, dist="weibull", data = df)
#summary(weibull)

loglogistic <- survreg(Surv(survival_in_days, status) ~ treatment + months_from_diagnosis + age_in_years + prior_chemotherapy + karnofsky_score + cell_type, dist="loglogistic", data = df)
#summary(loglogistic)

library(stargazer)
stargazer::stargazer(cox, exponential, weibull, loglogistic, title="Comparison of marginal effects using different models", type="text")
outfile = "Survival_models.html"
stargazer::stargazer(cox, exponential, weibull, loglogistic, title="Comparison of marginal effects using different survival models", out=outfile)
```

